Using CGM to Predict CFRD in the Envision Cohort

### 2025-07-01 Christine Chan/EnVision CF/cfrd_roc.qmd
simplified ROC for Zach

```{r data cleaning}
# Import cleaned dataset
load("./Data_Clean/envision_analysis_dataset.RData")
df <- final_df
# Limit to those with CGM data (based on mean glucose) and OGTT diagnosis
df <- df %>% filter(!is.na(mean_sensor))
# Convert CV units to match Scully et al.
df$cv <- df$cv * 100
```

- **All tables and analyses below treat CGM wears as independent**, but we will need to think about how to handle this in future analyses.

@@

```{r results='asis'}
t1 <- tableby(
  Diagnosis ~ redcap_data_access_group + sex + origin_race + ethnicity +
    pancreatic_status + CFTR + fev1 + fvc + bmi_perc + a1c_result,
  data = df
)
summary(t1, pfootnote = T)
```

@@

```{r results='asis'}
t2 <- tableby(
  Dysglycemia ~ redcap_data_access_group + sex + origin_race + ethnicity +
    pancreatic_status + CFTR + fev1 + fvc + bmi_perc + a1c_result,
  data = df
)
summary(t2, pfootnote = T)
```

@@

```{r results='asis'}
t3 <- tableby(
  CFRD ~ redcap_data_access_group + sex + origin_race + ethnicity +
    pancreatic_status + CFTR + fev1 + fvc + bmi_perc + a1c_result,
  data = df
)
summary(t3, pfootnote = T)
```

@@

```{r}
r <- roc(CFRD ~ mean_sensor, data = df, quiet = T)
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
gt(coords(r, "all", ret = "all")) %>% fmt_number(decimals = 3)
```

@@

```{r}
r <- roc(CFRD ~ percent_time_over_140, data = df, quiet = T)
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
gt(coords(r, "all", ret = "all")) %>% fmt_number(decimals = 3)
```

@@

```{r}
r <- roc(CFRD ~ percent_time_over_180, data = df, quiet = T)
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
gt(coords(r, "all", ret = "all")) %>% fmt_number(decimals = 3)
```

@@

```{r}
r <- roc(CFRD ~ gmi, data = df, quiet = T)
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
gt(coords(r, "all", ret = "all")) %>% fmt_number(decimals = 3)
```

@@

```{r}
r <- roc(CFRD ~ standard_deviation, data = df, quiet = T)
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
gt(coords(r, "all", ret = "all")) %>% fmt_number(decimals = 3)
```

@@

```{r}
r <- roc(CFRD ~ cv, data = df, quiet = T)
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
gt(coords(r, "all", ret = "all")) %>% fmt_number(decimals = 3)
```

@@

```{r}
r <- roc(CFRD ~ a1c_result, data = df, quiet = T)
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
gt(coords(r, "all", ret = "all")) %>% fmt_number(decimals = 3)
```

>>>>

```{r data cleaning}
set.seed(1017)
# Import cleaned dataset
load("./Data_Clean/envision_analysis_dataset.RData")
df <- final_df
# Limit to those with CGM data (based on mean glucose) and OGTT diagnosis
df <- df %>% filter(!is.na(mean_sensor), !is.na(Diagnosis))
# Convert CV units to match Scully et al.
df$cv <- df$cv * 100
# First CGM and OGTT pair only
df <- df %>%
  group_by(study_id) %>%
  slice_min(Date)
```

@@

```{r results='asis'}
t1 <- tableby(
  Diagnosis ~ redcap_data_access_group + age_visit + height + weight +
    bmi_perc + sex + origin_race + ethnicity + pancreatic_status + CFTR + fev1 +
    fvc + a1c_result + corrector_yes_no,
  data = df
)
summary(t1, pfootnote = T)
```

@@

```{r results='asis'}
t2 <- tableby(
  Dysglycemia ~ redcap_data_access_group + age_visit + height + weight +
    bmi_perc + sex + origin_race + ethnicity + pancreatic_status + CFTR + fev1 +
    fvc + a1c_result + corrector_yes_no,
  data = df
)
summary(t2, pfootnote = T)
```

@@

```{r results='asis'}
t3 <- tableby(
  CFRD ~ redcap_data_access_group + age_visit + height + weight +
    bmi_perc + sex + origin_race + ethnicity + pancreatic_status + CFTR + fev1 +
    fvc + a1c_result + corrector_yes_no,
  data = df
)
summary(t3, pfootnote = T)
```

@@

```{r}
r <- roc(CFRD ~ mean_sensor, data = df, quiet = T)
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
ci.auc(r)
# gt(data.frame(ci.coords(r, "best",
#   best.policy = "random",
#   ret = c(c("sensitivity", "specificity", "fp", "tp", "fn", "tn"))
# ))) %>% fmt_number(decimals = 3)
gt(coords(r, "best", ret = "all")) %>% fmt_number(decimals = 3)
```

@@

```{r}
r <- roc(CFRD ~ percent_time_over_140, data = df, quiet = T)
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
ci.auc(r)
# gt(data.frame(ci.coords(r, "best",
#   best.policy = "random",
#   ret = c(c("sensitivity", "specificity", "fp", "tp", "fn", "tn"))
# ))) %>% fmt_number(decimals = 3)
gt(coords(r, "best", ret = "all")) %>% fmt_number(decimals = 3)
```

@@

```{r}
r <- roc(CFRD ~ percent_time_over_180, data = df, quiet = T)
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
ci.auc(r)
# gt(data.frame(ci.coords(r, "best",
#   best.policy = "random",
#   ret = c(c("sensitivity", "specificity", "fp", "tp", "fn", "tn"))
# ))) %>% fmt_number(decimals = 3)
gt(coords(r, "best", ret = "all")) %>% fmt_number(decimals = 3)
```

@@

```{r}
r <- roc(CFRD ~ gmi, data = df, quiet = T)
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
ci.auc(r)
# gt(data.frame(ci.coords(r, "best",
#   best.policy = "random",
#   ret = c(c("sensitivity", "specificity", "fp", "tp", "fn", "tn"))
# ))) %>% fmt_number(decimals = 3)
gt(coords(r, "best", ret = "all")) %>% fmt_number(decimals = 3)
```

@@

```{r}
r <- roc(CFRD ~ standard_deviation, data = df, quiet = T)
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
ci.auc(r)
# gt(data.frame(ci.coords(r, "best",
#   best.policy = "random",
#   ret = c(c("sensitivity", "specificity", "fp", "tp", "fn", "tn"))
# ))) %>% fmt_number(decimals = 3)
gt(coords(r, "best", ret = "all")) %>% fmt_number(decimals = 3)
```

@@

```{r}
r <- roc(CFRD ~ cv, data = df, quiet = T)
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
ci.auc(r)
# gt(data.frame(ci.coords(r, "best",
#   best.policy = "random",
#   ret = c(c("sensitivity", "specificity", "fp", "tp", "fn", "tn"))
# ))) %>% fmt_number(decimals = 3)
gt(coords(r, "best", ret = "all")) %>% fmt_number(decimals = 3)
```

@@

```{r}
r <- roc(CFRD ~ a1c_result, data = df, quiet = T)
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
ci.auc(r)
# gt(data.frame(ci.coords(r, "best",
#   best.policy = "random",
#   ret = c(c("sensitivity", "specificity", "fp", "tp", "fn", "tn"))
# ))) %>% fmt_number(decimals = 3)
gt(coords(r, "best", ret = "all")) %>% fmt_number(decimals = 3)
```
